Automatically drawing vegetation maps using digital time-lapse cameras in alpine ecosystems

### 2022-06-06 paper/paper.tex
Updated paper

Previous studies have mainly depended on field observations, yet it is hard to cover broad areas in alpine regions due to poor accessibility and severe weather. Satellite, airborne, and Unmanned Aerial Vehicle (UAV) remote sensing methods seem to be alternatives. However, satellite imagery of alpine areas is rarely available due to cloud cover, and the spatial resolution is not enough to observe vegetation changes at the plant community scale. Airborne imagery can obtain high-resolution data, but its cost becomes a bottleneck for frequent monitoring. Although UAV methods have become a cost-effective tool for ecological monitoring (\cite{Baena2017PLOSONE}), operating UAVs in alpine regions is challenging due to the strong wind and harsh topology.

On the other hand, researchers have also utilized automated digital time-lapse cameras mounted on the ground for monitoring green-leaf phenologies in forests (\cite{Richardson2009EcolAppl}), grasslands (\cite{Browning2017RemSen}), and alpine meadows (\cite{IdeOguma2013EcolInfom}). Unlike satellite imagery, such cameras provide images free of clouds and atmospheric effects. Also, they can obtain high-resolution (i.e., sub-meter scale) and frequent (i.e., daily or hourly) images at a meager cost. These studies set some regions of interest (ROI) in the images and calculate the phenology index (e.g., excess greenness, \cite{Woebbecke1995ASAE}). Also, we can visually interpret the vegetation distribution and its changes from such time-lapse images. However, few studies have tackled this. This lack of studies seems to be because applying such time-lapse imagery in monitoring vegetation distribution has two technical challenges.

First, unlike satellites' multispectral sensors, ordinal digital cameras can only obtain three bands (Red, Green, and Blue), making it harder to classify the vegetation. Second, since digital time-lapse cameras are mounted on the ground, transforming these images into geospatial data (e.g., orthoimage) is challenging. In other words, even if we find a vegetation change, we cannot identify its geographic location and analyze it with other geographic data (e.g., topography). Treating ground-based time-lapse images as geographic data is essential to utilize them in conservation planning.

This study proposes an automated method for drawing vegetation maps and locating vegetation changes with a digital time-lapse camera by solving these two challenges. We solved the first challenge by utilizing temporal changes in leaf colors in vegetation classification. We also developed an accurate method for transforming a ground-based photograph into geographic data to tackle the second challenge. Finally, we show an example application of our method by detecting and localizing vegetation changes with an 8-years digital time-lapse imagery of the Japanese alps. We aim to use cheap but powerful digital time-lapse cameras in alpine ecosystem conservation.

@@

We used repeat photography data owned by National Institute for Environmental Studies, Japan (NIES). All the images are publically available on NIES' webpage (\url{https://db.cger.nies.go.jp/gem/ja/mountain/station.html?id=2}). In 2010, NIES installed the digital time-lapse camera (EOS 5D MK2, Canon Inc., 21 M pixels) on a mountain lodge Murodo-sanso (about 2350 m a.s.l., above the forest limit), located at the foot of Mt. Tateyama (3015 m a.s.l.), in the Nothern Japanese Alps. The camera takes one photograph per hour, from 6 a.m. to 7 p.m. . The camera's field of view (FOV) includes Mt. Tateyama, which ranges from about 2350 m a.s.l. to 3015 m a.s.l. in elevation. The area has a complex mosaic-like vegetation structure because of its topography, including rocks, cliffs, curls, and moraines. From April to November, the camera has observed the snowmelt and seasonal phenology of evergreen, deciduous dwarf trees (e.g., \emph{Pinus pumila}, \emph{Sorbus} sp.), dwarf bamboos (e.g., \emph{Sasa kurilensis}), and alpine herbaceous plants (e.g., \emph{Geum pentapetalum}, \emph{Nephrophyllidium crista-galli}). We used the images from late summer to late fall of 2012 and 2020.

@@

First, we selected images that are suitable for vegetation classification. We choose four days with good weather from late summer to late fall each year (9/19, 10/1, 10/11, 10/23 in 2010. 9/19, 10/2, 10/11, 10/23 in 2020). We used the images of this season because we can separate vegetation from the patterns of autumn foliage coloration. Also, we chose four images from 11 a.m. to 2 p.m. each day to avoid the effect of temporal noise such as shadows. Finally, we got 16 images for each year.

>>>>

Previous studies have mainly depended on field observations, yet it is hard to cover broad areas in alpine regions due to the poor accessibility and severe weather. Satellite, airborne, and Unmanned Aerial Vehicle (UAV) remote sensing methods seem to be alternatives. However, satellite imagery of alpine areas is rarely available due to cloud cover, and the spatial resolution is not enough to observe vegetation changes at the plant community scale. Airborne imagery can obtain high-resolution data, but its cost becomes a bottleneck for frequent monitoring. Although UAV methods have become a cost-effective tool for ecological monitoring (\cite{Baena2017PLOSONE}), operating UAVs in alpine regions is challenging due to the strong wind and harsh topology.

On the other hand, researchers have also utilized automated digital time-lapse cameras mounted on the ground for monitoring green-leaf phenologies in forests (\cite{Richardson2009EcolAppl}), grasslands (\cite{Browning2017RemSen}), and alpine meadows (\cite{IdeOguma2013EcolInfom}). Unlike satellite imagery, such cameras provide images free of clouds and atmospheric effects. Also, they can obtain high-resolution (i.e., sub-meter scale) and frequent (i.e., daily or hourly) images at a meager cost. These studies set some regions of interest (ROI) in the images and calculate the phenology index (e.g., excess greenness, \cite{Woebbecke1995ASAE}). Also, we can visually interpret the vegetation distribution and its changes from such time-lapse images. However, few studies have tackled this. This lack of studies seems to be because applying such repeat photography in monitoring vegetation distribution has two technical challenges.

First, unlike satellites' multispectral sensors, ordinal digital cameras can only obtain three bands (Red, Green, and Blue), making it harder to classify the vegetation. Second, since digital time-lapse cameras are mounted on the ground, transforming these images into geospatial data (e.g., orthoimage) is challenging. In other words, even if we find a vegetation change, we cannot quantitatively measure and analyze it as geographic data. It is essential to treat ground-based images as geographic data to utilize them in conservation planning.

This study proposes an automated method for drawing vegetation maps and locating and measuring vegetation changes with a digital time-lapse camera by solving these two challenges. We solved the first challenge by utilizing temporal changes in leaf colors in vegetation classification. We also developed an accurate method for transforming a ground-based photograph into geographic data to tackle the second challenge. Finally, we show an example application of our method by detecting and localizing vegetation changes with an 8-years digital time-lapse imagery of the Japanese alps. We aim to use cheap but powerful digital time-lapse cameras in alpine ecosystem conservation.

@@

We used repeat photography data owned by National Institute for Environmental Studies, Japan (NIES). All the images are publically available on NIES' webpage (\url{https://db.cger.nies.go.jp/gem/ja/mountain/station.html?id=2}). In 2010, NIES installed the digital time-lapse camera (EOS 5D MK2, Canon Inc., 21 M pixels) on a mountain lodge Murodo-sanso (about 2350 m a.s.l., above the forest limit), located at the foot of Mt. Tateyama (3015 m a.s.l.), in the Nothern Japanese Alps. The camera takes one photograph per hour, from 6 a.m. to 7 p.m. . The camera's field of view (FOV) includes Mt. Tateyama, which ranges from about 2350 m a.s.l. to 3015 m a.s.l. in elevation. The area has a complex mosaic-like vegetation structure because of its topography, including rocks, cliffs, curls, and moraines. From April to November, the camera has observed the snowmelt and seasonal phenology of evergreen, deciduous dwarf trees (e.g., \emph{Pinus pumila}, \emph{Sorbus} sp.), dwarf bamboos (e.g., \emph{Sasa kurilensis}), and alpine herbaceous plants (e.g., \emph{Geum pentapetalum}, \emph{Nephrophyllidium crista-galli}). We used the images from late summer to late fall of 2012 and 2020 to observe the distribution changes of the dwarf pine (\emph{Pinus pumila}): the dominant species in Japanese alps, and the invasive dwarf bamboo (\emph{Sasa kurilensis}).

@@

First, we selected images that are suitable for vegetation classification. We choose four days with good weather from late summer to late fall each year (9/19, 10/1, 10/11, 10/23 in 2010. 9/19, 10/2, 10/11, 10/23 in 2020). We used the images of this season because we can classify vegetation with the patterns of autumn foliage coloration. Also, we chose four images from 11 a.m. to 2 p.m. each day to avoid the effect of temporal noise such as shadows. Finally, we got 16 images for each year.
